---
title: "Figures"
author: "Kilian"
date: "2024-02-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This document summarize the figures which we have focused on, I leave everything in .Rmd and will maybe switch each function in a separate .R file.

# Libraries
```{r}
library(tidyverse)
library(cowplot)
library(RColorBrewer)
library(ggrepel)
library(patchwork)
```


# Data needed :
```{r}
working_dataset <- read.csv2(here("./data/dataset_V5.csv"))
atolls <- read.csv2(here("./data/atolls.csv"))
```


## Shapefile project, will be needed from 3th figure
```{r}
shp <- sf::st_read(here("./data/GEOGRAPHIE_DES_ATOLLS")) %>% #Not on Github for the moment, its in the GEOGRAPHIE_DES_ATOLLS file in the Biogeography Kilian
  sf::st_transform(crs = 4326) %>% 
  sf::st_make_valid()  %>% 
  dplyr::mutate(ILE = ifelse(OBJECTID %in% c(24733, 24734), "MOTU ONE (MAR)", ILE)) %>% #The shapefile does not distinct Motu One from Society and Marquesas archipelagos.
  dplyr::left_join(atolls, by = join_by(ILE)) # To add Archipelagos
```


# FIGURE 1 : Make the Sankey diagram
```{r}
# A connection data frame is a list of flows with intensity for each flow
links <- data.frame(
  source=c( "Animal : 2.98e+05", "Animal : 2.98e+05", "After 1950 : 2.77e+05", "After 1950 : 2.77e+05", "Presence : 2.77e+05", "Presence : 2.77e+05", "Records : 2.76e+05", "Records : 2.76e+05", "Unique : 1.69e+05", "Unique : 1.69e+05", "Habitat : 1.56e+05", "Habitat : 1.56e+05", "Habitat : 1.56e+05"), 
  
  target=c( "After 1950 : 2.77e+05", "Before 1950 : 2.1e+04", "Presence : 2.77e+05", "Absence data : 2e+02", "Records : 2.76e+05", "Invalid recording method : 2e+02", "True duplicates : 1.07e+05", "Unique : 1.69e+05","Habitat : 1.56e+05", "No habitat : 1.29e+04", "Marine", "Terrestrial", "Others"), 
  
  value=c(0.277 , 0.02, 0.276,  0.001, 0.275, 0.001, 0.107, 0.169, 0.156, 0.013, 0.11, 0.01, 0.036)
  )
 
# From these flows we need to create a node data frame: it lists every entities involved in the flow
nodes <- data.frame(
  name=c(as.character(links$source), 
  as.character(links$target)) %>% unique()
)
 
# With networkD3, connection must be provided using id, not using real name like in the links dataframe.. So we need to reformat it.
links$IDsource <- match(links$source, nodes$name)-1 
links$IDtarget <- match(links$target, nodes$name)-1

nodes$group <- as.factor(c( "A", "A", "A", "A", "A", "A", "B", "B", "B", "B", "B", "A", "A", "B")) 

links$group <- as.factor(c( "a","b", "a", "b", "a", "b", "b", "a", "a", "b", "a", "a", "b"))

my_color <- 'd3.scaleOrdinal() .domain(["a","b","A","B"]) .range(["#fdc30a","steelblue", "#daa520" ,"#4a9ac6"])'

# Make the Network
p <- networkD3::sankeyNetwork(Links = links, Nodes = nodes,
              Source = "IDsource", Target = "IDtarget",
              Value = "value", colourScale=my_color, LinkGroup="group", NodeGroup="group", NodeID = "name", 
              sinksRight=FALSE, fontSize = 13, width = 1000, height = 200, nodeWidth = 5, nodePadding = 20)
p


# save the widget
# library(htmlwidgets)
# saveWidget(p, file=paste0( getwd(), "/sankeyBasic1.html"))
```


# FIGURE 2 : Over and under representation.

```{r}
plot_ideal_sampling <- function(data, habitat, occurrence_threshold, n_tot, n_occ) {
  # Filtrer les données en fonction de l'habitat spécifié
  filtered_data <- data %>%
    filter(Habitat == habitat,
           !is.na(phylum)) 

  # Résumé des données en regroupant par archipel et niveau taxonomique
  summary_data <- filtered_data %>%
    mutate(
      nbocc = n(),
      idealsampling = n_occ * n_distinct(ScientificName) / n_tot,
      .by = c("class")) %>%
    mutate(Difference = nbocc - idealsampling, .by ="class") %>%
    filter(!is.na(nbocc) & nbocc >= occurrence_threshold) %>%
    mutate(bar_order = ifelse(Difference > 0, "Over-representation", "Under-representation"))
  
  # Créer le graphique
  ggplot(summary_data, aes(x = fct_reorder(phylum, Difference), y = asinh(Difference))) + 
    geom_bar(stat = "identity", aes(fill = bar_order), position = "stack") +
    scale_fill_manual(values = c("Over-representation" = "#1b9e77", "Under-representation" = "#d95f02"), name = "") +
    coord_flip() + 
    labs(title = "",
         x = "Class",  # Modifier le label de l'axe x
         y = "Occurrences") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) 
  #+ facet_wrap(~Archipel)
}

# Exemple d'utilisation de la fonction avec le niveau taxonomique "phylum"
mar <- plot_ideal_sampling(working_dataset, habitat = "Marine", occurrence_threshold = 100, n_tot = 5738, n_occ = 110890)

ter <- plot_ideal_sampling(working_dataset, habitat = "Terrestrial", occurrence_threshold = 10, n_tot = 980, n_occ = 9765)

# all <- plot_ideal_sampling(working_dataset, habitat = c("Terrestrial","Marine"), occurrence_threshold = 100, n_tot = 6675, n_occ = 120655)

# plot_grid <- cowplot::plot_grid(mar, ter, ncol = 2, align = "hv")

ggsave("figure2_ter.jpeg", ter, width = 10, height = 6, dpi = 600)
ggsave("figure2_mer.jpeg", mar, width = 10, height = 6, dpi = 600)
```

# Figure 3
```{r}
pal <- wesanderson::wes_palette("Zissou1", 100, type = "continuous")
colors <- RColorBrewer::brewer.pal(5, "Set1") 
labels = c("Australs", "Gambier", "Marquesas", "Society", "Tuamotu")

# PANEL A : Spatial distribution (you need to load shp file which is in GEOGRAPHIE DES ATOLLS)
fig3_panelA <- function(dataset, habitat, deg, shp_data) {
  
  dataset <- dataset %>%
    filter(Habitat == habitat) %>% 
    sf::st_as_sf(coords = c("decimalLongitude", "decimalLatitude"), crs = 4326)
  
  grid <- shp %>% 
    sf::st_make_grid(cellsize = c(deg, deg)) %>%
    sf::st_as_sf() %>% 
    mutate(grid_id = 1:n())%>%
    sf::st_join(dataset) %>%
    group_by(grid_id) %>%
    summarize(n_occ = n(),
              n_species = n_distinct(ScientificName))

  gg <- ggplot() +
    geom_sf(data = grid, aes(fill = log(n_species +1)), color = NA) +
    scale_fill_gradientn(colours = pal, limits = c(0,10)) +
    geom_sf(data = shp, alpha = 0.01, color = "white") +
    labs(title = "",
      x = "Longitude",
      y = "Latitude") +
    theme_bw() +
    theme(axis.text.x = element_text(color = "black", size = 12),  
        axis.text.y = element_text(color = "black", size = 12),
        axis.title.x = element_text(size = 15), axis.title.y = element_text(size = 15))
  
  return(gg)
  }

## PANEL B : Violin distribution 
fig3_panelB <- function(dataset, habitat) {
  
  dataset <- dataset %>%
    filter(Habitat == habitat) %>%
    dplyr::group_by(Archipel, ILE) %>%
    summarise(n_occ = n(),
              n_species = n_distinct(ScientificName))
    
    
  gg <- ggplot(data = dataset, aes(x = log10(n_species+1), y = Archipel, colour = Archipel)) +
    geom_point(show.legend = FALSE) +
    geom_violin(aes(fill = Archipel), alpha = 0.3, show.legend = FALSE) +
    labs(title = "",
         x = "Species (log)",
         y = "Archipelagos") +
    scale_color_manual(values = colors, labels = labels) +
    scale_fill_manual(values = colors, labels = labels) +
    scale_y_discrete(labels = labels) +
    ggrepel::geom_text_repel(aes(label = ILE), size = 3, color = "black") +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1, color = "black", size = 12),  
        axis.text.y = element_text(angle = 45, hjust = 1, color = "black", size = 12),
        axis.title.x = element_text(size = 15), axis.title.y = element_text(size = 15)) + 
    xlim(0,5)
  
  return(gg)
  }

```

# Figure 3 : FULL (Might take a little bit of time ~5min with n_grid = 50)
```{r}
a <- fig3_panelA(working_dataset, "Marine", 0.5, shp)
b <- fig3_panelA(working_dataset, "Terrestrial", 0.5, shp) + theme(axis.title.y = element_blank(), axis.text.y = element_blank())

occurrences_mar <- fig3_panelB(working_dataset, "Marine")
occurrences_terr <- fig3_panelB(working_dataset, "Terrestrial") + 
  theme(axis.title.y = element_blank(), axis.text.y = element_blank())

combined_plot <- cowplot::plot_grid(a, b, occurrences_mar, occurrences_terr, labels = c("Marine", "Terrestrial"), ncol = 2, nrow = 2) +
  theme(legend.position = 'top') 

ggsave("combined_plot.png", combined_plot, width = 8, height = 10, units = "cm", dpi = 600)

```


# Figure 4 (You can change this code, I've put mine to generate rarefaction curves based on occurrences,but I guess Laura will modify it)
```{r}
generate_rarefaction_curve <- function(dataset, threshold_percentage, habitat) {

  occurrences_by_species <- dataset %>%
    filter(Habitat == habitat) %>% #Archipel == archipel_name, 
    group_by(ScientificName, Archipel) %>%
    summarise(NombreOccurrences = n(),
              NombreEspeces = n_distinct(ScientificName))
  
  
  data_cumulative <- occurrences_by_species %>%
    arrange(NombreOccurrences) %>%
    group_by(Archipel) %>%
    mutate(NombreEspecesCumule = cumsum(NombreEspeces),
           NombreOccurrencesCumule = cumsum(NombreOccurrences))

  # Calcul du pourcentage d'espèces total
  total_species <- sum(data_cumulative$NombreEspeces)
  threshold <- total_species * threshold_percentage  # Seuil en fonction du pourcentage spécifié

  # Trouver la première occurrence où le nombre cumulé d'espèces dépasse le seuil de pourcentage spécifié
  first_occurrence_above_threshold <- data_cumulative %>%
    group_by(Archipel) %>%
    filter(NombreEspecesCumule >= threshold) %>% 
    slice(1)  # Récupérer la première ligne qui dépasse le seuil

  threshold_occurrence <- first_occurrence_above_threshold$NombreOccurrencesCumule  # Récupérer le nombre d'occurrences correspondant au seuil

  table_cumulative <- data.frame(NombreOccurrencesCumule = data_cumulative$NombreOccurrencesCumule,
                                 NombreEspecesCumule = data_cumulative$NombreEspecesCumule,
                                 Archipel = data_cumulative$Archipel)

  # Utilisation de ggplot pour créer le graphique
  p <- ggplot(table_cumulative, aes(x = NombreOccurrencesCumule, y = NombreEspecesCumule, color = Archipel)) +
    geom_line(show.legend = FALSE) +
    labs(title = "",
         x = "",
         y = "") +
    scale_color_manual(values = palette_archipel) +
    theme_bw() + xlim(0, 90000) + ylim(0,4500) 

  # Ajouter la ligne verticale au graphique

  
  #diviser les facets
  p <- p + facet_wrap(facets = ~Archipel, nrow = 1)
  
  p <- p + geom_vline(xintercept = threshold_occurrence, linetype = "dashed", color = "red") +
     annotate("text", x = 1.4 * threshold_occurrence, y = 0.95* max(data_cumulative$NombreEspecesCumule), label =  threshold_occurrence, vjust = -1)
  
  return(p)  # Retourner le graphique modifié
}

palette_archipel = c("#E41A1C", "#377EB8", "#4DAF4A", "#984EA3", "#FF7F00")

# global <- generate_rarefaction_curve(working_dataset, "All", "black", 0.75)
acc_cur_marine <- generate_rarefaction_curve(working_dataset,  0.75, "Marine") #"#984EA3",
acc_cur_terr <- generate_rarefaction_curve(working_dataset,  0.75, "Terrestrial")

cowplot::plot_grid(acc_cur_marine, acc_cur_terr, labels = c("Marine", "Terrestrial") ,nrow = 2)
```


# Figure 5 : Sampbias 
```{r}
data_sampbias <- read.csv2("C:/Users/Kilian/Desktop/Biogeography Kilian/Sample_bias_richness_q1.csv") # Sampbias package data stored in the file

fig5_plot <- function(data, sampbias, N_species, type) {
  
  # Fit linear model
  fit <- lm(log(N_species + 1) ~ sampbias, data = data)
  
  # Define color palette
  palette_archipel <- c("#E41A1C", "#377EB8", "#4DAF4A", "#984EA3", "#FF7F00")
  
  ggplot(data = data, aes(x = log(49 * sampbias + 1), y = log(N_species + 1))) +
    geom_point(aes(color = Archipel), show.legend = FALSE) +
    geom_smooth(method = lm, aes(color = Archipel, fill = Archipel), size = 1, linetype = "solid", alpha = 0.5, se = TRUE) +
    labs(title = "",
         x = "Expected occurrences (log)",
         y = "Observed occurrences (log)") +
    ggrepel::geom_text_repel(aes(label = ILE), size = 3, color = "black") +
    scale_color_manual(values = palette_archipel) +
    scale_fill_manual(values = palette_archipel) +
    theme(axis.text.x = element_text(size = 15), axis.text.y = element_text(size = 15), axis.title.y = element_text(size = 20)) +
    theme_bw() + xlim(0, 10) + ylim(0, 10) +
    facet_wrap(~ Archipel, scales = "free") +
    ggpubr::stat_regline_equation(label.x = 3, label.y = 10)

}

marine_sampbias_plot <- fig5_plot(data_sampbias, data_sampbias$Marine_sampbias, data_sampbias$n_occ_marine, "Marine") + 
  theme(axis.title.x = element_blank())
terrestrial_sampbias_plot <- fig5_plot(data_sampbias, data_sampbias$Terrestrial_sampbias, data_sampbias$n_occ_terrestrial, "Marine") + 
  theme(axis.title.x = element_blank())

sampbias_plot <- marine_sampbias_plot + terrestrial_sampbias_plot + plot_layout(guides = "collect", nrow = 2)

print(sampbias_plot)
```


```{r}
aus_sa <- fig4_plot_function_sampbias(ausq1, ausq1$Marine_sampbias, ausq1$n_occ_marine, "Marine", "#E41A1C", "Australs") + theme(axis.title.x = element_blank())
mar_sa <- fig4_plot_function_sampbias(marq1, marq1$Marine_sampbias, marq1$n_occ_marine, "Marine", "#4DAF4A", "Marquesas") + theme(axis.title.x = element_blank())
tua_sa <- fig4_plot_function_sampbias(tuaq1, tuaq1$Marine_sampbias, tuaq1$n_occ_marine, "Marine", "#FF7F00", "Tuamotu") 
gam_sa <- fig4_plot_function_sampbias(gamq1, gamq1$Marine_sampbias, gamq1$n_occ_marine, "Marine", "#377EB8", "Gambier") + theme(axis.title.x = element_blank())


soc_sa_t <- fig4_plot_function_sampbias(socq1, socq1$Terrestrial_sampbias, socq1$n_occ_terrestrial, "Terrestrial", "#984EA3", "Society") + theme(axis.title.y = element_blank()) + theme(axis.title.x = element_blank())
aus_sa_t <- fig4_plot_function_sampbias(ausq1, ausq1$Terrestrial_sampbias, ausq1$n_occ_terrestrial, "Terrestrial", "#E41A1C", "Australs") + theme(axis.title.y = element_blank()) + theme(axis.title.x = element_blank())
mar_sa_t <- fig4_plot_function_sampbias(marq1, marq1$Terrestrial_sampbias, marq1$n_occ_terrestrial, "Terrestrial", "#4DAF4A", "Marquesas") + theme(axis.title.y = element_blank()) + theme(axis.title.x = element_blank())
tua_sa_t <- fig4_plot_function_sampbias(tuaq1, tuaq1$Terrestrial_sampbias, tuaq1$n_occ_terrestrial, "Terrestrial", "#FF7F00", "Tuamotu") + theme(axis.title.y = element_blank())
gam_sa_t <- fig4_plot_function_sampbias(gamq1, gamq1$Terrestrial_sampbias, gamq1$n_occ_terrestrial, "Terrestrial", "#377EB8", "Gambier") + theme(axis.title.y = element_blank()) + theme(axis.title.x = element_blank())

library(patchwork)
sampbias_plot <- aus_sa + aus_sa_t + gam_sa + gam_sa_t + mar_sa + mar_sa_t + soc_sa + soc_sa_t + tua_sa + tua_sa_t + plot_layout(guides = "collect", ncol = 2, nrow = 5)
```

